% Define document class
\documentclass[twocolumn]{aastex631}
\usepackage{showyourwork}
\usepackage{amsmath}
\usepackage{bm}
\usepackage{listings}
\usepackage{natbib}
\usepackage{xcolor}


\definecolor{codegreen}{rgb}{0,0.6,0}
\definecolor{codegray}{rgb}{0.5,0.5,0.5}
\definecolor{codepurple}{rgb}{0.58,0,0.82}
\definecolor{backcolour}{rgb}{0.95,0.95,0.92}
\lstset{
    backgroundcolor=\color{white},   
    commentstyle=\color{codegreen},
    keywordstyle=\color{magenta},
    numberstyle=\tiny\color{codegray},
    stringstyle=\color{codepurple},
    basicstyle=\ttfamily\footnotesize,
    breakatwhitespace=false,
    breaklines=true,
    captionpos=b,
    keepspaces=true,
    numbers=left,
    numbersep=5pt,
    showspaces=false,
    showstringspaces=false,
    showtabs=false,
    tabsize=1
}

\definecolor{tedcommentcolor}{HTML}{e17701}
\newcommand{\TJ}[1]{\textcolor{tedcommentcolor}{#1}}

\DeclareMathOperator{\Span}{span}


% Begin!
\begin{document}

% Title
\title{Principal Component Analysis of Planetary Infrared Excess Data}

% Author list
\author{Ted Johnson}
\affiliation{UNLV}
\author{Avi Mandell}
\affiliation{GSFC}
\author{Rebecca Martin}
\affiliation{UNLV}

% Abstract with filler text
\begin{abstract}
    The Planetary Infrared Excess method to measure the thermal emission of an exoplanet is analogous to
    detecting the infrared excess of a circumstellar disk or low-mass companion -- except that the 
    exceptionally low contrast between planet and star require extreme precision and make the method
    easily susceptible to bias. We propose a new method to analyze PIE time-series data that naturally
    accounts for stellar variability and is agnostic to stellar spectral models. This method treats
    the observation as a vector space and isolates variability present at long wavelengths that is
    uncorrelated to short-wavelength variability.
\end{abstract}

% Main body with filler text
\section{Introduction}
\label{sec:intro}
I'll write this later.


\section{The Classical PIE Method}
\label{sec:pie}
Traditionally, the PIE method as described by \citet{stevenson2020} is as follows. A spectrograph measures a spectrum $\bm{f}$ from a planetary system with associated uncertainty $\bm{\sigma}$. Both are arrays, and we will write their elements as $f_i$ and $\sigma_i$, respectively. The spectrograph parameters are chosen so that $\bm{f}$ can be divided into two regions: $\bm{f}_{\rm NIR}$ is the short wavelength (near infrared) portion of the spectrum, where the star dominates, and the planet essentially contributes zero flux. $\bm{f}_{\rm MIR}$ is the long wavelength portion, where the flux from the planet is greater than the noise. The planetary system is unresolved, so
\begin{equation}
    \bm{f} = \bm{f}_{\rm s} + \bm{f}_{\rm p}
\end{equation}
where $\bm{f}_\text{s}$ is the stellar spectrum and $\bm{f}_\text{p}$ is the planetary spectrum.

The NIR and MIR portions of the spectrum are defined so that
\begin{equation}
    \bm{f}_{\rm NIR} = \bm{f}_{\rm s, NIR}
\end{equation}
and
\begin{equation}
    \bm{f}_{\rm MIR} = \bm{f}_{\rm s, MIR} + \bm{f}_{\rm p, MIR}
\end{equation}

In the classical PIE case we use a physics-based stellar model to produce a model spectrum $\tilde{\bm{f}}$. The parameters of that model are optimized so that $\bm{f}$ and $\tilde{\bm{f}}$ are approximately equal at short wavelengths. Due to noise and imperfections in the stellar model there is an error term, which the fitting process minimizes:
\begin{equation}
    \bm{\epsilon} \equiv \bm{f}_\text{s} - \tilde{\bm{f}}
\end{equation}
and since $\bm{f}_\text{NIR} = \bm{f}_\text{s, NIR}$, we can write
\begin{equation}
    \bm{\epsilon}_\text{NIR} = \bm{f}_\text{NIR} - \tilde{\bm{f}}_\text{NIR}
\end{equation}

The model $\tilde{\bm{f}}$ is then extrapolated to cover the full wavelength range (using the same model parameters as fit in the NIR). At long wavelengths, we subtract to find
\begin{equation}
    \label{eq:pie_result}
    \begin{array}{rcl}
        \bm{f}_\text{MIR} - \tilde{\bm{f}}_\text{MIR} & = & \bm{f}_{\rm s, MIR} + \bm{f}_{\rm p, MIR} - \tilde{\bm{f}}_\text{MIR} \\
        & = & \bm{\epsilon}_\text{MIR} + \bm{f}_{\rm p, MIR}
    \end{array}
\end{equation}

If $f_{\text{p},i} \gg \epsilon_i$, then we have successfully measured the planet's spectrum and can therefore infer its properties. However, recent high signal-to-noise studies with the James Webb Space Telescope (JWST) have shown that even our most sophisticated stellar models can only be expected to agree with observations to within a few percent \TJ{(ref? ask Avi what paper that was)}. Even in the case of a hot jupiter, where thermal emission could be on the order of $1\%$ of that of the star at 10 \textmu m, this error term dominates, and could introduce bias into any inferences made about the properties of the planet.

Use of modern observatories such as JWST mean that this classical PIE method is model-limited, and to improve our measurement of planetary flux we need to better subtract out the stellar contribution.

\section{The VPIE Method}
\label{sec:vpie}
The VPIE (Variable Planetary Infrared Excess) method aims to overcome the challenges described in Section \ref{sec:pie} by making a concession: we will only measure changes in the planet's flux. Just like in Section \ref{sec:pie}, we will start with an observation $\mathbf{F}$ with associated uncertainty $\mathbf{\Sigma}$, except that each of these is an $m \times n$ matrix where $m$ is the number of epochs and $n$ is the number of wavelength channels. We also define wavelength $\bm{\lambda}$ and time $\bm{t}$ such that ${\rm F}_{ij}$ is the flux at time $t_i$ and wavelength $\lambda_j$.

Our goal is the same: to compute some stellar model $\tilde{\mathbf{F}}$ to subtract from $\mathbf{F}$ in order to recover information from the planet. Again we do our fitting at short wavelengths only:
\begin{equation}
    \mathbf{F}_\text{NIR} - \tilde{\mathbf{F}}_\text{NIR} = \mathbf{E}_\text{NIR}
\end{equation}
where $\mathbf{E}$ is an error term defined as
\begin{equation}
    \label{eq:def_E}
    \mathbf{E} \equiv \mathbf{F}_\text{s} - \tilde{\mathbf{F}}_\text{s}
\end{equation}

Rather than compute $\tilde{\mathbf{F}}$ from some physical model, we will approximate as a matrix whose rows are linear combinations of a small subset of the rows of $\mathbf{F}$. We can choose a set of $q$ integers $s$, whose values are the indices of this ``spectral basis''. Let $\mathbf{B}$ be a $q\times n$ matrix
\begin{equation}
    B_{kj} = F_{s_kj} \, \text{for}\, k=1,\,2,\,...,\,q
\end{equation}

We can also define a coefficient matrix $\mathbf{A}$, which has shape $m\times q$. The model observation can be written
\begin{equation}
    \tilde{\mathbf{F}} = \mathbf{A}\mathbf{B}
\end{equation}

Choosing the best values for $\mathbf{A}$ and $s$ is discussed in Section \ref{??}. However, suppose we had found some optimal set of bases, described by $\hat{s}$, and then calculated $\mathbf{A}$ so that $\mathbf{E}_\text{NIR}$ is minimized. When computing the residual of the full wavelength range, we will introduce the matrix $\mathbf{\Delta}$, which is an additional error term which we attribute to the planet.
\begin{equation}
    \label{eq:measurable}
    \mathbf{F} - \tilde{\mathbf{F}} = \mathbf{E} + \mathbf{\Delta}
\end{equation}
As $\mathbf{F}$ is measurable and $\tilde{\mathbf{F}}$ can be calculated from it, the quantity $\mathbf{E} + \mathbf{\Delta}$ is measurable. We can decompose $\mathbf{F}$ and $\tilde{\mathbf{F}}$ to find its meaning. Recall
\begin{equation}
    \mathbf{F} = \mathbf{F}_\text{s} + \mathbf{F}_\text{p}
\end{equation}
so
\begin{equation}
    \mathbf{B} = \mathbf{B}_\text{s} + \mathbf{B}_\text{p}
\end{equation}
and $\tilde{\mathbf{F}}$ can be written
\begin{equation}
    \begin{array}{rccccc}
        \tilde{\mathbf{F}} & = & \mathbf{A}&(\mathbf{B}_\text{s} &+& \mathbf{B}_\text{p}) \\
        & = & & \tilde{\mathbf{F}}_\text{s} &+& \tilde{\mathbf{F}}_\text{p}
    \end{array}
\end{equation}
We can then rewrite Equation (\ref{eq:measurable}):
\begin{equation}
    \mathbf{E} + \mathbf{\Delta} = \mathbf{F}_\text{s} - \tilde{\mathbf{F}}_\text{s} + \mathbf{F}_\text{p} - \tilde{\mathbf{F}}_\text{p}
\end{equation}

Given the definition of $\mathbf{E}$ in Equation (\ref{eq:def_E}), we see that
\begin{equation}
    \mathbf{\Delta} =  \mathbf{F}_\text{p} - \tilde{\mathbf{F}}_\text{p}
\end{equation}

If $\mathbf{E}$ is small, then $\mathbf{F} - \tilde{\mathbf{F}}\approx \mathbf{\Delta}$, and we have measured some signal from the planet. Importantly, the expected value of $\mathbf{\Delta}$ can be calculated for any phase curve model, and therefore we can use our measured $\mathbf{\Delta}$ to infer properties of the planet.

In contrast to the method described in Section \ref{sec:pie}, our stellar error term $\mathbf{E}$ is not dominated by the limitations of a physics-based stellar model. Instead, there are two main sources of error: 1) complex variability: the time-dependent behavior of the star is too complicated to be represented by the $q$ basis vectors described by $\hat{s}$, and 2) measurement error, which in practice is mostly photon noise.

\TJ{(\#1 is unknown without real dataset, \#2 very doable)}




\section{Methods}
\label{sec:methods}

\subsection{Motivation}
We take a vector-based approach to time-series spectroscopy, representing spectra as vectors and observations as matrices.
\subsubsection{Spectroscopic Bases}
\label{subsubsec:bases}
A spectrograph discretizes a continuous spectrum into a series of wavelength bins. These spectral
elements are approximately linearly independent (though their values may be highly correlated). We can therefore
discuss a 1D spectrum as a vector of $n$ dimensions, where $n$ is the number of spectral elements. Let $S$ be a vector space of $n$ dimensions. A spectrum $\bm{f}$ can be written as
\begin{equation}
    \bm{f} = [f_1,\,f_2,\,...,\,f_j,\,...,\,f_n]
\end{equation}
where $f_j$ is a scalar value representing the flux of a single spectral channel.

Suppose the surface of a rotating star is made up of two regions:
\begin{enumerate}
    \item A quiet photosphere with intrinsic spectrum $\bm{f}_p$
    \item Low temperature spots, with intrinsic spectrum $\bm{f}_s$
\end{enumerate}

The fraction of each region visible to the observer is dependent on their viewing angle, and will change as the star rotates leading to a time-variable stellar spectrum. It can be written
\begin{equation}
    \label{eq:spec_basis_single}
    \bm{f}(t) = a_p(t) \bm{f}_p + a_s(t) \bm{f}_s
\end{equation}
where $a_p$ and $a_s$ are scalar functions of time. If this star is observed at times $\bm{t} = [t_1,\,t_2,\,t_3]$, then the resulting spectra will be 
\begin{equation}
    \label{eq:spec_basis_triple}
    \bm{f}(\bm{t})
    = \begin{bmatrix}
        a_p(t_1) \bm{f}_p + a_s(t_1) \bm{f}_s \\
        a_p(t_2) \bm{f}_p + a_s(t_2) \bm{f}_s \\
        a_p(t_3) \bm{f}_p + a_s(t_3) \bm{f}_s
    \end{bmatrix}
\end{equation}

Note that the right-hand sides of Equations (\ref{eq:spec_basis_single} \& \ref{eq:spec_basis_triple}) could be more concisely written if we define a basis $B = \{\bm{f}_p,\,\bm{f}_s\}$. In this basis, we can rewrite Equation (\ref{eq:spec_basis_single})
\begin{equation}
    [\bm{f}(t)]_{B} 
    = \begin{bmatrix}
        a_p(t),\,a_s(t)
    \end{bmatrix}
\end{equation}
And Equation \ref{eq:spec_basis_triple} is written
\begin{equation}
    [\bm{f}(\bm{t})]_{B} 
    = \begin{bmatrix}
        a_p(t_1) & a_s(t_1) \\
        a_p(t_2) & a_s(t_2) \\
        a_p(t_3) & a_s(t_3)
    \end{bmatrix}
\end{equation}


We generalize this to any astrophysical scene composed of objects whose respective scaling factors vary with time. For any time-variable scene whose variability can be written
\begin{equation}
    \bm{f}(t) = \sum_j a_j(t) \bm{f}_j
\end{equation}
a time series observation of $\bm{t} = [t_1,\,t_2,\,...\, t_i,\, ...,\,t_m]$ is then
\begin{equation}
    \label{eq:gen_spec_base_ts}
    \bm{f}(\bm{t})
    = \begin{bmatrix}
        \sum_j a_j(t_1) \bm{f}_j \\
        \sum_j a_j(t_2) \bm{f}_j \\
        ... \\
        \sum_j a_j(t_i) \bm{f}_j \\
        \dots \\
        \sum_j a_j(t_m) \bm{f}_j \\
    \end{bmatrix}
    = \mathcal{F}_{ij}
\end{equation}
where $\mathcal{F}_{ij}$ is a matrix of shape $m\times n$ whose rows are spectra and whose columns are light curves for a single spectral channel. We then define a basis $B = \{ \bm{f}_1,\,\bm{f}_2,\, ...,\, \bm{f}_j,\, ...,\, \bm{f}_p \}$ and use it to rewrite Equation (\ref{eq:gen_spec_base_ts})
\begin{equation}
    [\bm{f}(\bm{t})]_B 
    = \begin{bmatrix}
        a_1(t_1) & a_2(t_1) & ... & a_j(t_1) & ... & a_p(t_1) \\
        a_1(t_2) & a_2(t_2) & ... & a_j(t_2) & ... & a_p(t_2) \\
        \dots    & \dots    & ... & \dots    & ... & \dots    \\
        a_1(t_i) & a_2(t_i) & ... & a_j(t_i) & ... & a_p(t_i) \\
        \dots    & \dots    & ... & \dots    & ... & \dots    \\
        a_1(t_m) & a_2(t_m) & ... & a_j(t_m) & ... & a_p(t_m) \\
    \end{bmatrix}
\end{equation}
or, more concisely
\begin{equation}
    [\mathcal{F}_{ij}]_B = a_j(t_i)
\end{equation}

Where $[\mathcal{F}_{ij}]_B$ is an $m\times p$ matrix, where $p$ is the number of independent spectra that make up this variable scene. Herein we will assume that $p$ is small compared to $m$, that is, there are many more observed epochs than there are independent components in a model.

In practice, the basis $B$ is unknown. A spectroscopic time series resulting from an observation only contains scene-integrated spectra, and the actual spectra of the components ($\bm{f}_j$) are unknown. The time-series is written
\begin{equation}
    \label{eq:gen_obs_ts}
    \bm{f}(\bm{t})
    = \begin{bmatrix}
        \bm{f}(t_1) \\
        \bm{f}(t_2) \\
        ... \\
        \bm{f}(t_i) \\
        ... \\
        \bm{f}(t_m) \\
    \end{bmatrix}
\end{equation}
However, there is some set of $q \mid q \le p$ vectors $\bm{f}(t_i)$ that span this vector space. We define a set of indices $s$ to describe the epochs in which these vectors occur. We can then write the basis $B = \{\bm{f}(t_i)~\forall~i~\mid~i~\, \in\, s\}$ and use it to rewrite Equation (\ref{eq:gen_obs_ts}).
\begin{equation}
    [\bm(f)(\bm{t})]_B
    = \begin{bmatrix}
        b_i(\bm{t})\quad\forall ~i~\mid~i\,\in\,s
    \end{bmatrix}
\end{equation}
where $b_i(\bm{t})$ is the time-dependent coefficient on each of the basis vectors. The observation, initially described by an $m\times n$ matrix in Equation \ref{eq:gen_spec_base_ts}, is now described by two matrices: 1) a $q\times n$ matrix describing the basis, and 2) a $m\times q$ matrix describing the time-series in that basis. Any of our original vectors $\bm{f}(t_i)$ can be represented by a sum of the basis vectors
\begin{equation}
    \label{eq:single_epoch_sum}
    \bm{f}(t_i) = \sum_{i' \in s} b_{i'}(t_i)\bm{f}(t_{i'}) + \bm{\epsilon}(t_i)
\end{equation}
where $\bm{\epsilon}$ is an error term that in practice arises due to noise and deviations of an actual astrophysical scene from our strict assumptions. For a well-fitting model, $|\bm{\epsilon}|^2 \sim |\bm{\sigma}|^2$, where $\bm{\sigma}$ is the uncertainty on $\bm{f}$. Section \ref{subsec:implementation} will discuss methods to choose the set $s$ and the coefficients $b_{i'}(t_i)$ such that $\bm{\epsilon}^2$ is minimized.

\subsubsection{Application to PIE}

A prototypical PIE observation consists of an unresolved astrophysical scene consisting of a star and a planet observed across two bands, the near
infrared (NIR) and mid infrared (MIR). The star -- being much hotter than the planet -- contributes flux to both bands. The planet is not hot enough
to contribute to the NIR band, and only produces measurable flux in the MIR. We can write the flux in epoch $i$
\begin{equation}
    \label{eq:obs_composition}
    \bm{f}(t_i) = \bm{f}_s(t_i) + \bm{f}_p(t_i)
\end{equation}
for time-dependent stellar spectrum $\bm{f}_s(t_i)$ and time-dependent planetary spectrum $\bm{f}_p(t_i)$. In the NIR (i.e. where the planet contributes 0 flux), this is written
\begin{equation}
    \bm{f}^{\text{NIR}}(t_i) = \bm{f}^\text{NIR}_s(t_i)
\end{equation}

Considering this star-only portion of the spectrum, we can write it similar to Equation (\ref{eq:single_epoch_sum}).
\begin{equation}
    \begin{array}{ccl}
        \bm{f}^{\text{NIR}}(t_i) &=& \sum_{i' \in s} b_{i'}(t_i) \bm{f}^\text{NIR}(t_i') + \bm{\epsilon}^\text{NIR}(t_i) \\
        &=& \tilde{\bm{f}}^{\text{NIR}}(t_i) + \bm{\epsilon}^\text{NIR}(t_i)
    \end{array}
\end{equation}
where $s$ and $b_{i'}(t_i)$ are calculated using the NIR data only, and $\tilde{\bm{f}}$ represents the reconstruction of $\bm{f}$ using the basis $B$. We then use these NIR-only bases and coefficients to compute a reconstruction of the whole spectrum.
\begin{equation}
    \begin{array}{ccl}
        \bm{f}(t_i) &=& \sum_{i' \in s} b_{i'}(t_i) \bm{f}(t_i') + \bm{\epsilon}(t_i) + \bm{\delta}(t_i) \\
        &=& \tilde{\bm{f}}(t_i) + \bm{\epsilon}(t_i) + \bm{\delta}(t_i)
    \end{array}
\end{equation}
Where $\bm{\delta}$ is an additional error term due to the uncorrelated variability of the planet. By subtracting $\tilde{\bm{f}}(t_i)$ from $\bm{f}(t_i)$ we are able to measure the quantity $\bm{\epsilon}(t_i) + \bm{\delta}(t_i)$. We can interpret this quantity by substituting in Equation (\ref{eq:obs_composition})
\begin{equation}
    \begin{array}{ccl}
        \bm{f}_s(t_i) + \bm{f}_p(t_i) &=& \sum_{i' \in s} b_{i'}(t_i) \bm{f}_s(t_i')  \\
        & & + \sum_{i' \in s} b_{i'}(t_i) \bm{f}_p(t_i') \\
        & & + \bm{\epsilon}(t_i) + \bm{\delta}(t_i)
    \end{array}
\end{equation}
or
\begin{equation}
    \bm{\epsilon}(t_i) + \bm{\delta}(t_i) = \bm{f}_s(t_i) - \tilde{\bm{f}_s}(t_i) + \bm{f}_p(t_i) - \tilde{\bm{f}_p}(t_i)
\end{equation}
Since $\bm{\epsilon}(t_i)$ is defined as $\bm{f}_s(t_i) - \tilde{\bm{f}_s}(t_i)$, we are left with a definition for $\bm{\delta}(t_i)$:
\begin{equation}
    \begin{array}{ccl}
        \bm{\delta}(t_i) &=& \bm{f}_p(t_i) - \tilde{\bm{f}_p}(t_i) \\
        &=& \bm{f}_p(t_i) - \sum_{i'\in s} b_{i'}(t_i) \bm{f}_p(t_i')
    \end{array}
\end{equation}

The quantity $\bm{\epsilon}(t_i) + \bm{\delta}(t_i)$ can be measured, however, only $\bm{\delta}(t_i)$ provides any information about the planet's spectrum. Therefore, $\bm{\epsilon}(t_i)$ is not important so long as it is small. For simplicity, we will assume that $\bm{\delta}(t_i) \gg \bm{\epsilon}(t_i)$ in the MIR, the only wavelength band in which $\bm{\delta}$ is defined. This assumption allows for a direct measurement of $\bm{f}_p(t_i) - \tilde{\bm{f}_p}(t_i)$. To compute $\tilde{\bm{f}_p}(t_i)$ from some planetary phase curve model $\bm{f}_p(t)$ requires the set of basis indices $s$ as well as the coefficients $b_{i}(t)$, both of which are computed using the NIR data only and are independent of the planetary model. Therefore, any forward model of a planetary phase curve can be used to predict the observable error term $\bm{\delta}(t)$. Section \ref{subsec:implementation} will use this to compare the measured $\bm{\delta}(t)$ to the expected $\bm{\delta}(t)$ given by a grid of planetary models in order to retrieve the characteristics of the observed planet.

\subsection{Implementation}
\label{subsec:implementation}
The main challenge of this method is in determining the best set of indices $s$ to use as the basis, and from that, the coefficients $b_i(t)$. The latter can be accomplished using a form of least-squares minimuization.

\subsubsection{Determining the coefficients $b_i(t)$}
\label{subsubsec:coeffs}

Given some guess of indices $s'$, we construct a basis $B = \{ \bm{f}(t_i)\quad \forall\quad i \mid i \in s' \}$. Some reconstruction at epoch $t=t_i$ is written
\begin{equation}
    \tilde{\bm{f}}(t_i) = \sum_{i'\in s'} b_{i'}(t_i)\bm{f}(t_{i'})
\end{equation}
which can be written as a product of a row and column vector
\begin{equation}
    \label{eq:mat_reconstruct}
    \begin{array}{ccl}
        \tilde{\bm{f}}(t_i) & = &
        \begin{bmatrix}
            b_a(t_i) & b_b(t_i) & ...
        \end{bmatrix}
        \begin{bmatrix}
            \bm{f}(t_a) \\ \bm{f}(t_b) \\ ...
        \end{bmatrix} \\
        & & \forall\quad\{a,b,...\} \in s' \\
        & = & \bm{b}(t_i)\,\mathcal{B}
    \end{array}
\end{equation}
where $\mathcal{B}$ is the $q\times n$ basis matrix described in Section \ref{subsubsec:bases}, and $\bm{b}(t_i)$ is simply a row vector of $n$ coefficients. We want to choose coefficients that minimize the square of the error term $\bm{\epsilon}$, written
\begin{equation}
    |\bm{\epsilon}(t_i)\,\mathcal{W}(t_i)|^2 = | \bm{f}(t_i)\,\mathcal{W}(t_i) - \bm{b}(t_i)\, \mathcal{B} \,\mathcal{W}(t_i) |^2
\end{equation}
where $\mathcal{W}$ is a diagonal $n\times n$ matrix made up of the observed uncertainties (i.e. $\mathcal{W}_{jj}(t_i) = 1/\bm{\sigma}_j(t_i)$).

This is equivalent to finding the least-squares solution to the transposed equation:
\begin{equation}
   \mathcal{W}(t_i)\, \mathcal{B}^\top \bm{b}(t_i)^\top = \mathcal{W}(t_i)\,\bm{f}(t_i)^\top
\end{equation}
We show in Appendix \ref{ap:lsq} that the solution is
\begin{equation}
    \bm{b}(t_i) = \bm{f}(t_i)\,\mathcal{W}^2(t_i)\, \mathcal{B}^\top\,(\mathcal{B}\,\mathcal{W}^2(t_i)\,\mathcal{B}^\top)^{-1}
\end{equation}
We will define a matrix $\mathcal{M}(t_i)$ in order to rewrite Equation (\ref{eq:mat_reconstruct}).
\begin{equation}
    \mathcal{M}(t_i) = \mathcal{W}^2(t_i)\, \mathcal{B}^\top\,(\mathcal{B}\,\mathcal{W}^2(t_i)\,\mathcal{B}^\top)^{-1}
\end{equation}
\begin{equation}
    \label{eq:estimate}
    \tilde{\bm{f}}(t_i) = \bm{f}(t_i)\,\mathcal{M}(t_i)\mathcal{B}
\end{equation}
If $\bm{\sigma}$ does not change much with time, it may be computationally efficient to replace $\mathcal{W}(t)$ with $\langle {\mathcal{W}(t)}\rangle$ so that $\mathcal{M}$ only needs to be computed once per guess for set $s'$. Equation \ref{eq:estimate} gives the best approximation of $\bm{f}(t_i)$ given some set of basis vectors defined by $s'$. Next is to optimize $s'$ to give the best reconstruction.

\subsubsection{Determining the Basis}
\label{subsubsec:get_basis}
Our goal is to find $s$, a set of indices that describe the optimal basis of spectra. To do this we will test sets $s'$ and compare them using the Bayes Information Criterion (BIC, ref?), a model selection criterion for which smaller values are preferred:
\begin{equation}
    \text{BIC} = q\ln{(mn)} - 2\ln{(\hat{L})}
\end{equation}
where $\hat{L}$ is the maximized likelihood function. This is the probability of observing $\bm{f}(\bm{t})$ given the prediction given by reconstruction $\tilde{\bm{f}}(\bm{t})$ and measurement uncertainties $\bm{\sigma}(\bm{t})$.
\begin{equation}
    \hat{L} = \prod_{i=1}^m \prod_{j=1}^n \frac{1}{\sqrt{2\pi \sigma_j^2(t_i)}} \exp{\left (-\frac{(f_j(t_i) - \tilde{f}_j(t_i))^2}{2\sigma_j^2(t_i)}\right )}
\end{equation}
$\tilde{\bm{f}}(\bm{t})$ has already been computed using the algorithm in Section \ref{subsubsec:coeffs}, so this likelihood is maximized. To compute the BIC, we need $\ln{\hat{L}}$:
\begin{equation}
        \ln{\hat{L}} = \sum_{i=1}^m \sum_{j=1}^n -\frac{1}{2} \left (\ln{(2\pi\sigma_j^2(t_i))} + \frac{(f_j(t_i) - \tilde{f}_j(t_i))^2}{\sigma_j^2(t_i)}\right )
\end{equation}
Written with vector-wise operations for simplicity, the BIC is
\begin{equation}
    \label{eq:bic2}
    \text{BIC} = q\ln{(mn)} + \sum \left (\ln{(2\pi\bm{\sigma}^2(\bm{t}))} + \frac{(\bm{f}(\bm{t})- \tilde{\bm{f}}(\bm{t}))^2}{\bm{\sigma}^2(\bm{t})}\right )
\end{equation}

The only values in Equation(\ref{eq:bic2}) that change for any given $s'$ are $q$, the number of elements in $s'$, and $\tilde{\bm{f}}(\bm{t})$, which is computed given $s'$ in Section \ref{subsubsec:coeffs}. Therefore, there exists some function $\text{BIC}(s')$ that can be used to compare any choice of $s'$ to all other choices of $s'$.

All that is left now is to test various values of $s'$, and choose the one for which BIC is minimized. There are many ways to do this. The simplest is to start at low values of $q$ and step through each possible value of $s'$ sequentially, finding a local minimum for each value of $q$. When increasing $q$ no longer leads to decreased BIC, the search is stopped and the previous minimum is returned as the solution.

\begin{lstlisting}[
    label={ls:soln1}, caption={Minimizing BIC though an exhaustive search at low $q$}
]
minima = {}
for q in 0..m-1:
    sets = combos(0..m-1, q)
    vals = []
    for s in sets:
        vals.append(bic(s))
    imin = argmin(vals)
    if q > 1:
        if vals[imin] < minima[q-1]["val"]:
            minima[q] = {
                "val":vals[imin],
                "s":sets[imin]
            }
        else:
            return minima[q-1]["s"]
    else:
            minima[q] = {
                "val":vals[imin],
                "s":sets[imin]
            }
\end{lstlisting}

This method, shown as pseudocode in Listing \ref{ls:soln1} is exhaustive, but quickly becomes computationally expensive. Given a time series of 100 spectra and a basis of size 2 (i.e., $m=100$, $q=2$, respectively) there are 4950 possible values of $s'$. However, if instead $q=4$, there are $3\,921\,225$ possible values of $s'$.

A less expensive method is to find local minimum with $q=1$, and then add a single index at a time as $q$ is increased. This will not necessarily find the global minimum of the BIC, but searches higher values of $q$ much faster. A pseudocode implementation is shown in Listing \ref{ls:sequential}.

\begin{lstlisting}[
    label={ls:sequential},caption={Minimizing BIC by searching for the next best index for each value of $q$}
]
s = []
vals = []
for _ in 1..m-1:
    _is = []
    _vals = []
    for i in 1..m-1 if i not in s:
        _s = s + [i]
        _is.append(i)
        _vals.append(bic(_s))
    imin = argmin(_vals)
    s.append(_is[imin])
    vals.append(_vals[imin])
q_best = argmin(vals) + 1
s = s[:q_best-1]
return s
\end{lstlisting}

\appendix

This is mostly so that I don't forget things during the process of writing. They may not be that important.
\section{Math}
\subsection{Least Squares Algorithm}
\label{ap:lsq}
Recall from Section \ref{subsubsec:coeffs} that we would like to solve the equation:
\begin{equation}
    \bm{f}\,\mathcal{W} = \bm{b}\,\mathcal{B}\,\mathcal{W}
\end{equation}
Where $\bm{f}$ is $1\times n$, $\bm{b}$ is $1\times q$, $\mathcal{W}$ is $n\times n$, and $\mathcal{B}$ is $q\times n$. We transpose both sides to get a more familiar form.
\begin{equation}
    \mathcal{W}\,\mathcal{B}^\top\,\bm{b}^\top = \mathcal{W}\,\bm{f}^\top
\end{equation}
The solution to this equation is known from least squares
\begin{equation}
    \bm{b}^\top = \mathcal{Q}\,\mathcal{W}\,\bm{f}^\top
\end{equation}
where $\mathcal{Q}$ is the Moore-Penrose pseudoinverse Matrix (ref?), defined by
\begin{equation}
    \mathcal{Q} = ((\mathcal{W}\,\mathcal{B}^\top)^\top\,\mathcal{W}\,\mathcal{B}^\top)^{-1}\,(\mathcal{W}\,\mathcal{B}^\top)^\top
\end{equation}
This leaves us with an expression for $\bm{b}$
\begin{equation}
    \begin{array}{ccl}
        \bm{b} &=& \bm{f}\,\mathcal{W}\,\mathcal{Q}^\top \\
        &=& \bm{f}\,\mathcal{W}\, ( (\mathcal{BWWB}^\top)^{-1} \mathcal{BW})^\top\\
        &=& \bm{f}\,\mathcal{W}\,\mathcal{W}\,\mathcal{B}^\top ((\mathcal{BWWB}^\top)^\top)^{-1}\\
        &=& \bm{f}\,\mathcal{W}^2\, \mathcal{B}^\top\,(\mathcal{B}\,\mathcal{W}^2\,\mathcal{B}^\top)^{-1} \\
    \end{array}
\end{equation}

\bibliography{syw,pie}

\end{document}
